---
title: "Pilot ECDRE 16S comparison"
author: "Jason Stajich"
date: "November 8, 2021"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(root.dir = normalizePath(".."))
min_read_count=1000
```

## Comparing the DNA extraction protocol and 16S profile

```{r warning=FALSE, message=FALSE}
library(ape)
library(vegan)
library(dplyr)
library(scales)
library(grid)
library(reshape2)
library(phyloseq)
library(magrittr)
library(ggplot2)
library(ggpubr)
library(plyr)
library(data.table)
library(tidyr)
library(tidyverse)
```
###STEP2: Import Mapping file (metadate file)
1.Check mapping file before import to R, R doesn’t seem to like sample name to start with number or contain “-” in sample name. If you get error in this step, you should check file name first.

2.First column of first row should not start with #, R will not read the first row that starts with #

```{r}
metatable = file.path(knitr::opts_chunk$get("root.dir"),"Pilot_16S.mapping_file.txt")
meta = read.table(metatable,header=TRUE,row.names=1,sep=",",stringsAsFactors=FALSE)
rownames(meta) <- gsub('-','.',rownames(meta))
```

If need to keep or remove some samples
```{r}
meta <- meta[which(meta$DemuxReads > min_read_count),]
```

###STEP3: Check if your metadata file has been import successfully and correctly

The output will show a table of your metadata file (mapping file).

*If you do not have header, you might have started your first row with #

```{r warning=FALSE}
head(meta)
```

###STEP4: Construct sample_data-class using imported metadata

```{r}
sampleData <- sample_data(meta)
```

###STEP5: Import OTU table

OTU table from Mojave 16S data is “MC2017FC.otu_table.txt”.
MC-Mojave Crust-2017-FC-Fungi completed

```{r}
otutable <- file.path(knitr::opts_chunk$get("root.dir"), "Pilot_16S.otu_table.txt")
otus <- read.table(otutable,header=T,sep="\t",row.names=1)
otumat <- as(as.matrix(otus), "matrix")
OTU = otu_table(otumat, taxa_are_rows = TRUE)
```

Check imported OTU table

```{r warning=FALSE}
head(OTU)
```

###STEP6: Import taxonomy table
Taxonmy table generated from AMPtk need to be rearranged using following script.

“perl rdp_taxonmy2mat.pl<Input_taxonmy.txt>Output_taxonomy.txt”

rdp_taxonomy2mat.pl was created by Professor Jason E. Stajich

```{r}
taxin = file.path(knitr::opts_chunk$get("root.dir"),"Pilot_16S.cluster.taxonomy.fix.txt")
taxmat <- read.table(taxin, header=T,sep="\t",row.names=1)
taxmat <- as(as.matrix(taxmat),"matrix")
TAX = tax_table(taxmat)
```

###STEP7: Import phylogenetic tree
Phylogenetic tree can also be include for further phylogenetic analysis.

```{r warning=FALSE}
treefile = file.path(knitr::opts_chunk$get("root.dir"),"Pilot_16S.cluster.tree.phy")
tree = read.tree(treefile)
```

###STEP8: Construct Phyloseq object
To construct phyloseq object, otu table, taxonomy table, and sampleData are required. Phylogenetic tree can be included, but it is not necessary for constructing phyloseq object.
Construct Phyloseq object called "Physeq"

```{r warning=FALSE}
physeq = phyloseq(OTU,TAX,sampleData,tree)
```

Check phyloseq object
This should indicate that your physeq is a "phyloseq-class experiment-level object""

```{r warning=FALSE}
physeq
```

###STEP9: Remove singletons
Remove any OTUs that present only one time.

```{r }
physeq.prune = prune_taxa(taxa_sums(physeq) > 1, physeq)
```

```{r warning=FALSE}
physeq.prune
```

###STEP10: Plot read counts to check dataset
Check read counts: any samples that have very low reads should be removed.
[Ref](http://evomics.org/wp-content/uploads/2016/01/phyloseq-Lab-01-Answers.html)

```{r}
readcount = data.table(as(sample_data(physeq.prune), "data.frame"),
                 TotalReads = sample_sums(physeq.prune), 
                 keep.rownames = TRUE)
setnames(readcount, "rn", "SampleID")
#For plotting, use command below.
#SeqDepth = ggplot(readcount, aes(TotalReads)) + geom_histogram() + ggtitle("Sequencing Depth")
```

TotalReads of all the samples can be in this table (select only SampleID and TotalReads columns).
In order to check samples with low number of reads, "order()" can be used to sort "TotalReads" column.
In this dataset, N55.Rhizo has very low number of reads, so will will filter this sample out using the next minimum number of reads.
```{r}
readcount = readcount[order(readcount$TotalReads), c("SampleID", "TotalReads")]
```

```{r}
head(readcount)
```

```{r warning=FALSE}
set.seed(1)
physeq.prune.rarefy = rarefy_even_depth(physeq.prune, sample.size = 1500, replace = FALSE, trimOTUs = FALSE)
physeq.prune.rarefy
```

###STEP11.1: Plot Alpha diversity by Crust_type
Alpha diversity can be Chao1, Observed, Shannon, Simpson
This plot include statistical analysis using "stat_compare_means" with "method = anova"


###STEP12.1 Taxonomic composition

```{r}
BactOrder = as.character(get_taxa_unique(physeq.prune.rarefy, "Order"))
BactOrder = BactOrder[complete.cases(BactOrder)]
BactOrder
```

```{r fig.align='center'}
plot_richness(physeq.prune.rarefy, x = "ExtractByTmt", color = "ExtractByTmt", measures = c("Shannon")) + 
  geom_boxplot() +
  theme_bw() + ggtitle("Alpha diversity plot by sample (Shannon)") + 
  theme(plot.title = element_text(hjust = 0.5)) + 
  geom_point(size=5, alpha=0.8)
```

###STEP12:Taxonomy Barplot

Taxonomy barplot by sample (Class level)

```{r fig.align='center'}
plot_bar(physeq.prune.rarefy, x = "Sample", y = "Abundance", fill ="Class") + 
  geom_bar(aes(color=Class, fill=Class), stat="identity", position="stack") +
  ggtitle("Taxonomy barplot by sample (Phylum Class)") + 
  theme(plot.title = element_text(hjust = 0.5))
```

###Beta Diversity

```{r}
physeq.prune.rarefy.ps.ord <- ordinate(physeq.prune.rarefy, "PCoA", "unifrac")
```

```{r}
plot_ordination(physeq.prune.rarefy, physeq.prune.rarefy.ps.ord, type = "samples", color = "Extraction",shape="Tissue")  + ggtitle("Bacterial Beta Diversity (PCoA) by Owner")
```

```{r fig.height=5, fig.width=8, fig.align="center"}
taxcom_treatment = ggplot(data = psmelt(physeq.prune.rarefy), mapping = aes_string(x = "ExtractByTmt",
                                                                                   y = "Abundance", 
                                                                                   fill = "Phylum" )) + 
  geom_bar(stat="identity", position="fill") + 
  ggtitle("Bacteria from pilot study")+
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
                     axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)
                     ) +
  theme(plot.title = element_text(hjust = 0.5))
```

```{r}
print(taxcom_treatment)
```

Save plot
```{r}
pdf("../Figures/Plot_name.pdf", width = 6, height = 3.8)
print(taxcom_treatment)
dev.off()
```

###Plot Alpha Diversity (Species richness)

```{r}
physeq.prune.rarefy.plot.richness = plot_richness(physeq.prune.rarefy, x="Treatment", color=("Extraction"), measures=c("Observed")) + geom_boxplot(lwd=0.5) + ggtitle("Alpha Diversity by Treatment") + stat_compare_means(method = "t.test", label.x.npc = c("right")) + theme_bw() + scale_colour_brewer(palette="Set1") + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
physeq.prune.rarefy.plot.richness + theme(plot.title = element_text(hjust = 0.5))
```


